---
title: "Surface Water Chem"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fs)
library(glue)
library(httr)
library(jsonlite)
library(lubridate)
library(vroom)
```

```{r}
source('R/update-aos-files.R')
lake_river_sites <- c('BARC', 'SUGG', 'CRAM', 'LIRO', 
                      'PRPO', 'PRLA', 'TOOK', 'FLNT', 
                      'BLWA', 'TOMB', 'BLUE')
aq_site_ids <- readr::read_lines('aq_site_ids.txt')
```

## Update surface water chemistry data

```{r}
chem_box_dir = '~/Box/data/NEON/NEON_chem-surfacewater'
myglob = 'externalLabData'
chem_data_id = 'DP1.20093.001' # surface water chem
mysite = 'PRLA'

# # update_swchem_files(aq_site_ids[5])
lake_river_sites %>% 
  purrr::walk(~update_aos_files(chem_box_dir, 
                                chem_data_id, 
                                mysite = .x, 
                                myglob = 'externalLabData'))
lake_river_sites %>% 
  purrr::walk(~update_aos_files(chem_box_dir, 
                                chem_data_id, 
                                mysite = .x, 
                                myglob = 'fieldSuperParent'))
lake_river_sites %>% 
  purrr::walk(~update_aos_files(chem_box_dir, 
                                chem_data_id, 
                                mysite = .x, 
                                myglob = 'swc_fieldData'))
```

## Update phytoplankter data

```{r}
phyto_dir = '~/Box/data/NEON/NEON_chem-peri-ses-phyto'
phyto_data_id = 'DP1.20163.001' # phytoplankton chem
update_aos_files(phyto_dir, phyto_data_id, mysite = 'FLNT', myglob = 'alg_fieldData')

aq_site_ids %>% 
  purrr::walk(~update_aos_files(phyto_dir, 
                                phyto_data_id, 
                                mysite = .x, 
                                myglob = 'alg_fieldData'))
aq_site_ids %>% 
  purrr::walk(~update_aos_files(phyto_dir, 
                                phyto_data_id, 
                                mysite = .x, 
                                myglob = 'algaeExternalLabDataPerSample'))

```


## Update phytoplankter taxonomy data

```{r}
taxo_dir <- '~/Box/data/NEON/NEON_peri-ses-phyto'
taxo_data_id <- 'DP1.20166.001' # phytoplankton taxonomy
myglob <- 'alg_taxonomyProcessed'

lake_river_sites %>% 
  purrr::walk(~update_aos_files(taxo_dir, 
                                taxo_data_id, 
                                mysite = .x, 
                                myglob = 'alg_taxonomyProcessed'))

```

## Combine chlorophyll data

```{r}
phyto_dir2 <- '~/Box/data/NEON-processed/chl-a'

# join the field and lab data files 
# by site-months
# mysite = 'POSE'
join_field_and_lab <- function(mysite){
  glue::glue('{phyto_dir}/{mysite}') %>% fs::dir_create()
  glue::glue('{phyto_dir2}/{mysite}') %>% fs::dir_create()
  lab_files <- fs::dir_ls(glue('{phyto_dir}/{mysite}'), glob = glue::glue("*algaeExternalLabDataPerSample*"))
  field_files <- fs::dir_ls(glue('{phyto_dir}/{mysite}'), glob = glue::glue("*alg_fieldData*"))
  
  lab_months_have <- basename(lab_files) %>% 
    stringr::str_split(pattern = '\\.', simplify = FALSE) %>%
    purrr::map_chr(~.x[[8]])
  
  processed_months <- fs::dir_ls(glue::glue('{phyto_dir2}/{mysite}')) %>% 
    basename() %>% tools::file_path_sans_ext() %>%
    stringr::str_split('_') %>% purrr::map_chr(~.x[[2]])
  
  months_to_join <- lab_months_have[!lab_months_have %in% processed_months]
  # my_month <- months_to_join[1]
  join_month <- function(my_month){
    lab_df <- grep(my_month, lab_files, value = TRUE) %>% 
      readr::read_csv(col_types = readr::cols_only(domainID = 'c',
                                            siteID = 'c',
                                            sampleType = 'c',
                                            namedLocation = 'c',
                                            collectDate = 'D',
                                            sampleID = 'c',
                                            sampleCondition = 'c',
                                            replicate = 'd',
                                            analyte = 'c',
                                            analyteConcentration = 'd',
                                            plantAlgaeLabUnits = 'c',
                                            externalLabDataQF = 'c')) %>% 
      # readr::read_csv(col_types = 'ccccDccccDccddddcdccccc') %>%
      dplyr::filter(sampleType %in% c('chlorophyll/pheophytin')) %>%
      dplyr::select(domainID, siteID, namedLocation, collectDate,
                    sampleID, sampleCondition, replicate,
                    analyte, analyteConcentration, plantAlgaeLabUnits,
                    externalLabDataQF)
    field_cols <- 'cccccdddddcTcccccccccccccccccc'
    field_df <- grep(my_month, field_files, value = TRUE) %>% 
      readr::read_csv(col_types = field_cols) %>% 
      dplyr::mutate(time_hms = hms::as_hms(collectDate),
             collect_date = lubridate::as_date(collectDate)) %>%
      dplyr::select(namedLocation, aquaticSiteType, collect_date,
                    time_hms, parentSampleID, habitatType,
                    algalSampleType, phytoDepth1,
                    phytoDepth2, phytoDepth3, substratumSizeClass)
    
    # make table between sampleId and parentSampleID
    # based on field table
    my_lookups <- field_df$parentSampleID %>%
      purrr::set_names(field_df$parentSampleID) %>%
      purrr::map(~grep(.x, lab_df$sampleID, value = TRUE)) %>%
      purrr::map_df(~as_tibble(.x), .id = 'parentSampleID') %>%
      dplyr::rename(sampleID = value) %>% 
      dplyr::distinct()

    lab_join_field <- lab_df %>% 
      dplyr::left_join(my_lookups, by = 'sampleID') %>%
      dplyr::left_join(field_df, by = c('parentSampleID', 'namedLocation')) 
    
    lab_join_field %>% readr::write_csv(glue::glue('{phyto_dir2}/{mysite}/{mysite}_{my_month}.csv'))
    message(glue::glue('Saved chl a lab data for {mysite} {my_month}'))
    # return(lab_join_field)
  }
  
  months_to_join %>% purrr::walk(~join_month(.x))
  # all_lab_join_field %>%
    # write_csv(glue('{phyto_dir2}/{mysite}_phyto-chem.csv'))
}
# join_field_and_lab('POSE')
# aq_site_ids %>% purrr::walk(~join_field_and_lab(.x))
lake_river_sites %>% purrr::walk(~join_field_and_lab(.x))
```

## Organize algal taxonomy

acceptedTaxonID

Morphospecies Cyanobacterial spheres is "UNKALG" taxonID


for diatoms: 

* algalAnalysisMethod = diatom slide
* algalType = Bacillariophyceae
* algalParameter = valves
* division = Chrysophyta	
* class = Bacillariophyceae

```{r}
mysite = 'PRPO'
taxo_dir2 <- '~/Box/data/NEON-processed/alg-taxa'

 
  glue::glue('{taxo_dir2}/{mysite}') %>% fs::dir_create()
  lab_files <- glue::glue('{taxo_dir}/{mysite}') %>% fs::dir_ls()

  lab_months_have <- basename(lab_files) %>% 
    stringr::str_split(pattern = '\\.', simplify = FALSE) %>%
    purrr::map_chr(~.x[[8]])
  
  processed_months <- fs::dir_ls(glue::glue('{taxo_dir2}/{mysite}')) %>% 
    basename() %>% tools::file_path_sans_ext() %>%
    stringr::str_split('_') %>% purrr::map_chr(~.x[[2]])
  
  months_to_process <- lab_months_have[!lab_months_have %in% processed_months]
  
  my_month <- months_to_process[14]

  lab_df <- grep(my_month, lab_files, value = TRUE) %>% 
      readr::read_csv(col_types = readr::cols_only(domainID = 'c',
                                            siteID = 'c',
                                            namedLocation = 'c',
                                            collectDate = 'T',
                                            sampleID = 'c',
                                            sampleCondition = 'c',
                                            algalType = 'c',
                                            algalParameter = 'c',
                                            algalParameterValue = 'd',
                                            acceptedTaxonID = 'c',
                                            scientificName = 'c',
                                            morphospeciesID = 'c',
                                            division = 'c',
                                            class = 'c',
                                            order = 'c',
                                            family = 'c',
                                            genus = 'c',
                                            taxonRank = 'c')) %>%
    dplyr::mutate(collectDate = lubridate::as_date(collectDate)) %>%
    dplyr::filter(namedLocation %in% pref_locs_df$namedLocation)
  
  lab_df_top10 <- lab_df %>% dplyr::group_by(sampleID) %>%
    dplyr::arrange(-algalParameterValue) %>%
    dplyr::slice_head(n = 10)
```

